---
title: "RV144 Aids trials"
author: "instr"
date: "April 13, 2016"
output: html_document
---


# Thailand Aids study
```{r include=FALSE}
require(tigerstats)
```

## Introduction

In the late 1980s in Thailand, there was a dramatic increase in the prevalence of infection with the human immunodeficiency virus type 1 (HIV-1) in sentinel surveillance cohorts.1-3 Initially, these groups consisted of injection-drug users and commercial sex workers; they were subsequently expanded to include persons in the general population. By 1995, the overall seroprevalence of HIV-1 reached a peak of 3.7% among conscripts in the Royal Thai Army and of 12.5% among conscripts from Northern Thailand.2,4,5 The Thai Ministry of Public Health responded with an effective HIV-prevention campaign, and the number of new HIV-1 infections per year decreased from an estimated 143,000 in 1990 to 14,000 in 2007.2,4,6-9 The persistence of new infection despite these measures led public health officials to conclude that an HIV vaccine, within the context of a broader HIV-prevention program, was needed for better control of the epidemic.



## Methods and data

Study Design and Population
This study was a community-based, randomized, multicenter, double-blind, placebo-controlled efficacy trial of the prime–boost combination of vaccines containing ALVAC-HIV (vCP1521) (Sanofi Pasteur) and AIDSVAX B/E (Global Solutions for Infectious Diseases). For details regarding the vaccines and placebo, see the Supplementary Appendix, available with the full text of this article at NEJM.org. The study was designed to evaluate two coprimary end points: the prevention of HIV-1 infection and the effect of vaccination on the early viral load after infection. The trial was conducted through facilities of the Thai Ministry of Public Health in Rayong and Chon Buri provinces. From September 2003 through December 2005, a total of 16,402 volunteers were enrolled.
Thai men and women who were between the ages of 18 and 30 years and who were not infected with HIV were recruited from the community without regard to HIV risk (i.e., community risk). Written informed consent was obtained from all volunteers, who were required to pass a written test of understanding. Women were counseled to practice effective contraception until 3 months after the last vaccination; pregnant and breast-feeding women were excluded.

###Study Design and Population
This study was a community-based, randomized, multicenter, double-blind, placebo-controlled efficacy trial of the prime–boost combination of vaccines containing ALVAC-HIV (vCP1521) (Sanofi Pasteur) and AIDSVAX B/E (Global Solutions for Infectious Diseases). For details regarding the vaccines and placebo, see the Supplementary Appendix, available with the full text of this article at NEJM.org. The study was designed to evaluate two coprimary end points: the prevention of HIV-1 infection and the effect of vaccination on the early viral load after infection. The trial was conducted through facilities of the Thai Ministry of Public Health in Rayong and Chon Buri provinces. From September 2003 through December 2005, a total of 16,402 volunteers were enrolled.
Thai men and women who were between the ages of 18 and 30 years and who were not infected with HIV were recruited from the community without regard to HIV risk (i.e., community risk). Written informed consent was obtained from all volunteers, who were required to pass a written test of understanding. Women were counseled to practice effective contraception until 3 months after the last vaccination; pregnant and breast-feeding women were excluded.

###Primary End Points
HIV-1 Infection
HIV-1 infection was diagnosed in 132 subjects (56 in the vaccine group and 76 in the placebo group) during 52,985 person-years of follow-up in the intention-to-treat analysis, in 86 subjects (36 in the vaccine group and 50 in the placebo group) during 36,720 person-years of follow-up in the per-protocol analysis, and in 125 subjects (51 in the vaccine group and 74 in the placebo group) during 52,985 person-years of follow-up in the modified intention-to-treat analysis. One subject in the placebo group who was identified by hospital record as being seropositive for HIV after dying from Pneumocystis jirovecii pneumonia was included in the analysis before the unblinding of the study. This diagnosis of HIV-1 infection was the only one that occurred outside planned procedures.
We create the following data set


```{r}
results <- read.csv(file = "ThaiAidsTrial.csv")
trialtbl=xtabs(~results+treatment,data=results)

```


## Results
<!-- You are to add descriptive comments to explain to the reader what the plot is showing !-->

### Graphical results
<!--Display graphical evidence as to the effectiveness of the vaccine with explanations!-->
```{r}
barchartGC(trialtbl,type = "percent")
```

### Numerical Results

<!--Display numerical evidence as to the effectiveness of the vaccine with explanations !--> 

```{r}
colPerc(trialtbl)
rowPerc(trialtbl)
summary(results)
```




## Inferential results

<!--Display inferential evidence as to the effectiveness of the vaccine with explanations and interpretations!-->

Here is the  situation - there are 125 people that ended up with aids. We will simulate with 125 red balls. That leaves 16277 people without aids, we will simulate that with 16277 white balls. Now we "dream up" a measure of "oddness" that we can apply to any splitting of our set of balls into 2 equal sized groups.  We expect the redballs to be equally mixed in our two samples so we square the difference from 62.5 for each split and call that our "oddness measure" for that split. Lets figure out what our observed oddness number is (74-62.5)^2  + (51-62.5)^2 which is 132.25 + 132.25 = 264.5 .  Now we randomly sample our balls into two equal size groups and each time we compute our "oddness measure".  After doing this 10000 times we see where our oddness measure fits in with the 10000. When it ends up in the tail we tend to reject the null hypothesis.


```{r}
recordTrials = numeric()
recordTrials2 = numeric()
bucket=c(rep(1,times=125),rep(0,times=16277))
for (i in 1:10000){
  zsamp=sample(bucket,size=16402)
  front=sum(zsamp[1:8201])
  back=sum(zsamp[8202:16402])
  recordTrials[i] = (front-back)^2 
  recordTrials2[i] = abs(front-back)
}
```
```{r}
ourecdf <- ecdf(recordTrials)
ourecdf2 <- ecdf(recordTrials2)
ourecdf((74-51)^2)
quantile(recordTrials,.95)
ourecdf2(abs(74-51))
quantile(recordTrials2,.95)
```


## Conclusions

<Summarise your results as to the effectiveness of the vaccine with explanations that could be understood by your mom>

We set up to see if we would get results similiar to that of our clinical trial under the conditions assumed by the null hypothesis. After 10000 simulations results computed by our "oddness measure" there were more extreme results only about 3 percent of the time.  This leads us to tenatively reject the null hypothesis. Notice also that monitone increasing functions give the same results when they are the measures of "oddness". The emperical cumulative distribution function is similiar to the p-variant we have already seen. Just like pnorm(x) will give the probablility of getting something less than x from a standard normal distribution so does ourecdf(x) give the probability that our newly created distribution is less than what ever value of x we plug in. The other function qnorm(p) is handled with the quantiles function in combination with the data that generates our emperical distribution.
See the examples using recordTrials and recordtrials2 above.